---
title: "Leaf_Stomatal_Guard_cell_models"
author: "Jiaxin Wang"
date: "2023-03-31"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data

```{r, echo=FALSE}
library(readxl)
```

```{r}
guardcell <- read_excel("D:/STOMATA/Stomata pictures from Heidi/images_for_guard cell estimation/model/guardcell.xlsx", 
                        sheet = "Sheet1")
```


## Plot and simple linear regression for guard cell width

```{r}
plot(guardcell$w~guardcell$g_width, ylab = "Guardcell width (pixels)", xlab = "Whole_stomata width (pixels)")
```


Fit a multivariate linear regression model to predict guard cell width


```{r}
lm01 <- lm(g_width~w+h+w2+h2, data = guardcell)
summary(lm01)
```

Get residuals and convert them to absolute values and use 1/absolute residuals as the weights to fit a weighted multivariate linear regression model


```{r}
guardcell$residuals1 <- abs(residuals(lm01))

lm01.wt <- lm(g_width~w+h+w2+h2, data = guardcell, weights = 1/residuals1)
summary(lm01.wt)
```

Check the assumptions

```{r}
plot(lm01.wt)
```



Fit a multivariate linear regression model to predict guard cell length

```{r}
lm02 <- lm(g_length~w+h+w2+h2, data = guardcell)
summary(lm02)
```


Get residuals and convert them to absolute values and use 1/absolute residuals as the weights to fit a weighted multivariate linear regression model

```{r}
guardcell$residuals2 <- abs(residuals(lm02))

lm02.wt <- lm(g_length~w+h+w2+h2, data = guardcell, weights = 1/residuals2)
summary(lm02.wt)
```


Check the assumptions

```{r}
plot(lm02.wt)
```


Generate predicted width and length of leaf stomatal guard cell

```{r}
guardcell$predict_width <- predict(lm01.wt)

guardcell$predict_length <- predict(lm02.wt)
```


## Plot predicted vs ImageJ measured leaf stomatal guard cell width and length

```{r}
plot(guardcell$predict_width,guardcell$g_width, xlab = "Predicted guard cell width", ylab = "Measured guard cell width")
abline(a=0, b=1, col = "blue")

plot(guardcell$predict_length,guardcell$g_length, xlab = "Predicted guard cell length", ylab = "Measured guard cell length")
abline(a=0, b=1, col = "blue")
```


Define Bias and RMSE function to calculate Bias and RMSE 


```{r}
bias <-  function(obs, pred, n){
  bias <-  round(sum(obs - pred)/length(obs), digit=4)
  bias.perc <-  round(bias/mean(obs)*100, digit=4)
  out  <-  c(bias,bias.perc)
  return(out)
}

RMSE <-  function(obs, pred, n){
  RMSE <-  round(sqrt(sum((obs - pred)^2)/length(obs)), digit=4)
  RMSE.perc <-  round(RMSE/mean(obs)*100, digit=4)
  out <-  c(RMSE,RMSE.perc)
  return(out)
}
```


```{r}
RMSE(guardcell$g_width,guardcell$predict_width,length(guardcell$g_width))

RMSE(guardcell$g_length,guardcell$predict_length,length(guardcell$g_length))
```











